The TMB R package

### 2014-10-17 _knitr-posts/2014-10-17-tmb.Rmd
Tweak TMB post and publish it

---
layout: post
title: The TMB R package
subtitle: Life beyond ADMB
published: false
---

I'd heard whispers of this new R package [TMB](https://github.com/kaskr/adcomp) and finally spent some time playing with it in the last few days. It's incredible — fast like ADMB (faster?), has built-in parallel functionality, has other advanced [statistical niceties](https://github.com/kaskr/adcomp/wiki), and it's built around R. What more could you ask for? (Well better documenation and the option to switch on Bayesian MCMC like you can in ADMB would be nice.)

The basic idea is to write a C++ template file and then compile and run it with some R functions. If you've used ADMB then the format will look similar. If you haven't, it's not too hard to pick up. It helps to [know a bit about C++](http://adv-r.had.co.nz/Rcpp.html), but for the most part it shouldn't be that hard for anyone familiar with R to pick it up.

@@

You can install the TMB package from GitHub:

@@

```{r tmb1, fig.height=5}
set.seed(1)
x <- runif(50, 1, 40)
y <- rnorm(50, mean = 2.5 + 0.3 * x, sd = exp(0.6))
plot(x, y)
```

@@

We can check the estimates against the true values above (`a = 2.5, b = 0.3, log_sigma = 0.6`):

@@

Which is exactly the same as we'd get from `stats::lm()`:

@@

```{r tmb-test1, results='hide', message=FALSE, warning=FALSE, fig.height=8}
sim_gomp(N = 5000, sigma_obs = 0.1)
```

>>>>

---
layout: post
title: The TMB R package
subtitle: Life beyond ADMB?
published: true
---

I'd heard whispers of this new R package [TMB](https://github.com/kaskr/adcomp) and finally spent some time playing with it in the last few days. It's incredible — fast like ADMB (faster?), has built-in parallel functionality, has other advanced [statistical niceties](https://github.com/kaskr/adcomp/wiki), and it's built around R. What more could you ask for? (Well better documenation and the option to switch on Bayesian MCMC like you can in ADMB would be nice.) This post just represents some of my initial steps as I learned and played with the package.

The basic idea with TMB is to write a C++ template file and then compile and run it with some R functions. If you've used ADMB then the format will look similar. If you haven't, it's not too hard to pick up. It helps to [know a bit about C++](http://adv-r.had.co.nz/Rcpp.html), but for the most part it shouldn't be that hard for anyone familiar with R to pick it up.

@@

You can install the TMB package from GitHub. You'll need a C++ compiler installed. If you're on Windows then you'll also need Rtools. See specific instructions on the [TMB GitHub page](https://github.com/kaskr/adcomp).

@@

```{r tmb1, fig.height=5}
set.seed(123)
x <- runif(20, 1, 10)
y <- rnorm(20, mean = 1.8 + 2.4 * x, sd = exp(0.3))
plot(x, y)
```

@@

We can check the estimates against the true values above (`a = 1.8, b = 2.4, log_sigma = 0.3`):

@@

Which is about the same as we'd get from `stats::lm()`:

@@

```{r tmb-test1, results='hide', message=FALSE, warning=FALSE, fig.height=8}
sim_gomp(N = 5000, sigma_obs = 0.1, sigma_proc = 0.2, b = 0.7, a = 1.4)
```
